---
title: 'Visual Analytic: Exercise UNI'
author: "Quinhas Fernandes"
date: "1/16/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(uwIntroStats)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(readr)
library(foreign)
library(haven)

load(file = url('https://github.com/EvansDataScience/data/raw/master/eduwa.rda'))

link="https://github.com/quinhasf/pubpol-599/raw/master/ape_analysis.dta"

chw_fp <- read_dta(url(link))

library(ggplot2)
```


```{r include=FALSE, results='hold'}

absoluteT=table(eduwa$LocaleType,exclude = 'nothing')

ToPlot=prop.table(absoluteT)

names(ToPlot)[5]='Unknown'

ToPlot=ToPlot*100 # preparing labels

ToPlot
```


```{r include=FALSE, results='hold'}
titleText='Where are Public Schools located in WA in 2019?'
sourceText='Source: US Department of Education'
```


```{r echo=FALSE, fig.show='hide'}
plotLabels=paste0(round(ToPlot,2),'%')

par(mgp=c(1,0.5,0)) 
location=barplot(ToPlot,
         col = 'gray',
         border=NA,
         main=titleText,
         xlim=c(0,50), 
         xlab = '(in %)',
         cex.names=0.6,
         cex.lab=0.6,
         las=2,
         horiz = T,
         xaxt="n") # no x-axis, so I customize it below...

text(x=ToPlot,y=location,labels=plotLabels,pos = 4,cex = 0.7) 
title(sub=sourceText, adj=0,cex.sub=0.7,line = 3) 

#reference line
abline(v=25,lty=3,lwd=3)


# customizing tick values
newXvalues<-c(0,10,25,40,50) # you just want to show this on the axis
axis(side=1, 
     at=newXvalues, 
     labels = newXvalues,
     cex.axis=0.8)
```


```{r include=FALSE, results='hold'}
tableFreq=as.data.frame(ToPlot)
names(tableFreq)=c("locale","pct")
#you have:
tableFreq
```


```{r fig.show='hide', include=FALSE}
library(ggplot2)

#base GGPLOT2 starts with a "base", telling WHAT VARIABLES TO PLOT
base= ggplot(data = tableFreq, 
             aes(x = locale,
                 y = pct)) 

plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') # y is just what it is!

plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)

plot3 = plot2 + geom_hline(yintercept = 25, #where
                           linetype="dashed", 
                           size=1.5, #thickness
                           alpha=0.5) #transparency
library(scales)

# customize Y axis
plot4 = plot3 + scale_y_continuous(breaks=c(0,10, 25,40,50),
                                 limits = c(0, 50),
                                 labels=scales::unit_format(suffix = '%')) 
plot5 = plot4 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.caption = element_text(hjust = 0), # default was 1
                    plot.title = element_text(hjust = 0.5))

plot6 = plot5 + geom_text(aes(y = pct ,
                            label = paste0(round(pct,2), '%')),
                        vjust=1, # if flipping 'hjust'
                        size = 3)

plot6
```


```{r include=FALSE, results='hold'}
plot6 + coord_flip()
```

This is what I want:


```{r echo=FALSE, results='hold'}
ToPlotOrd=sort(ToPlot)

par(mgp=c(1,0.5,0)) # distance label, tickText,tick
location=barplot(ToPlotOrd,
         col = 'gray',
         border=NA,
         main=titleText,
         xlim=c(0,50), 
         xlab = '(in %)',
         horiz = T,
         cex.names = 0.7,
         cex.lab=0.6,
         xaxt="n") # no x-axis, so I customize it below...

text(x=ToPlotOrd,y=location,labels=plotLabels,pos = 2,cex = 0.7) 
title(sub=sourceText, adj=0,cex.sub=0.7,line = 3) 

# reference line
abline(v=25,lty=3,lwd=3)

# customizong tick values
xtick<-c(0,10,25,40,50)
axis(side=1, at=xtick, labels = xtick,cex.axis=0.8)
```


But in GGPLOT… (exercise 1Use ggplot to show a bar plot ordered by share size.)

```{r include=FALSE, results='hold'}
tableFreq[order(-tableFreq$pct),]
```


```{r echo=TRUE, fig.show='hold'}

#base GGPLOT2 starts with a "base", telling WHAT VARIABLES TO PLOT
base= ggplot(data = tableFreq, 
              aes(x = reorder(locale, pct),
                 y = pct))

plot1 = base + geom_bar(fill ="gray",
                        stat = 'identity') # y is just what it is!

plot2 = plot1 + labs(title=titleText,
                     x =NULL, 
                     y = NULL,
                     caption = sourceText)

plot3 = plot2 + geom_hline(yintercept = 25, #where
                           linetype="dashed", 
                           size=1.5, #thickness
                           alpha=0.5) #transparency
library(scales)

# customize Y axis
plot4 = plot3 + scale_y_continuous(breaks=c(0,10, 25,40,50),
                                 limits = c(0, 50),
                                 labels=scales::unit_format(suffix = '%')) 
plot5 = plot4 + theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.caption = element_text(hjust = 0), # default was 1
                    plot.title = element_text(hjust = 0.5))

plot6 = plot5 + geom_text(aes(y = pct ,
                            label = paste0(round(pct,2), '%')),
                        vjust=1, # if flipping 'hjust'
                        size = 3)

plot7 = plot6 + coord_flip()

plot7
```

Exercise Lolliplot


```{r ggplot, include=FALSE}
chw_pf1 <- data.frame(chw_fp[c('province', 'ape_contrib')])

tableFreq=as.data.frame(chw_pf1)

names(tableFreq)=c("province","ape_contrib")

tableFreqO=tableFreq[order(tableFreq$ape_contrib),]

tableFreqO$gap=tableFreqO$ape_contrib-10
```


```{r lolliplot, echo=TRUE, fig.show='hold'}

tableFreqO$Target=ifelse(tableFreqO$gap>0,"Above Target","Below Target")

base = ggplot(tableFreqO, aes(province,gap,color=Target,
                              label = round(gap,1))) 
lolliplot1=base + geom_segment(aes(y =0, 
                                   x = province, 
                                   yend = gap, 
                                   xend = province)) 
lolliplot2=lolliplot1 + geom_point()

lolliplot3= lolliplot2 + scale_x_discrete(limits=tableFreqO$province) 

lollitplot4 = lolliplot3 + geom_text(nudge_x=0.3, nudge_y=0.1,show.legend = F) +
              labs(title = "Mozambique Community Health Workers contribuition on Family Planning",
                   subtitle = "2017 province level gap on CHW contribuition",
                    x ="Province", 
                    y = "% points of the FP GAP",
                    caption = "Fig.1: Represents each province contribuition to achieve the 10% target (centered at 0) in 2017 (Gap analysis).
Provinces are ploted from low to high perfomance.

Source:HIS") +
            theme(panel.background = element_rect(fill = "gray98",
                                                    colour = "black"),
                    plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5),
                    plot.subtitle = element_text(hjust=0.5),
                    legend.box.just = c("right","center"), 
                    axis.text.x = element_text(size=7, angle = 45, vjust = 1, hjust=1)) +
              geom_hline(yintercept=0,
                    linetype="dashed",
                    color = "black",
                    size= 0.9,
                    alpha= 0.8)


lollitplot4 
```


Class Exercise: Make sure our box plot follows the same design approach and include all the elements as in the bar plot for nominal data

```{r ordinalTable, include=FALSE}
table(eduwa$High.Grade,exclude = 'nothing')
```

```{r tableAsDF, include=FALSE}
frqTabO=as.data.frame(prop.table(table(eduwa$High.Grade)))
names(frqTabO)=c('grade','pct')
frqTabO
```

```{r echo=TRUE}
box1 = ggplot(eduwa, aes(y=as.numeric(High.Grade))) 
box1 = box1 + geom_boxplot(varwidth=T, fill="grey50") + coord_flip() 

ordLabels= levels(eduwa$High.Grade)

box2 = box1 + scale_y_continuous(labels=ordLabels,breaks=1:15)

box3 = box2 + 
    labs(title="2017 King County Students Grade Distribution", 
         subtitle="Box-plot of students high-grade distribution",
         caption="Source: Seattle Public Schools",
         y="Schools Grade") +
   theme(panel.background = element_rect(fill = "white",
                                                    colour = "grey50"),
                    plot.caption = element_text(hjust = 0), 
                    plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust=0.5)) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
box3
```

Class Exercise: Make a histogram for the variable FREE LUNCH, and make sure it has all the right elements, and get rid of unnecessary elements.

```{r echo=TRUE}

eduwa3=eduwa[complete.cases(eduwa$Reduced.Lunch),]

#ggplot
base= ggplot(eduwa3,aes(x = Free.Lunch))  
hist1= base + geom_histogram(binwidth = 100,boundary=0, fill="gray20") #changing width
hist2= hist1 + stat_bin(binwidth = 100, aes(label=..count..), 
                  geom = "text",boundary = 0,vjust=-0.5) +
                  labs(title="2017 Schools distribuition of Free Lunch", 
                           subtitle="Histogram of frequencies",
                       caption = "Source: Seattle Public Schools",
                       x= "Free Lunch",
                       y="Number of meals (counts)") +
                       xlim(c(0, 1800)) +
                 scale_x_continuous(breaks = c(0, 400, 800, 1200, 1600))
hist2
```

Class Exercise: Do some research and make a histogram and a density plot using ggplot for the variable we just used above.

```{r removeMissing2, include=FALSE}

eduwa4=eduwa[complete.cases(eduwa$Student.Teacher.Ratio),]

meanValues=tapply(eduwa4$Student.Teacher.Ratio, 
                  eduwa4$County, 
                  mean)
eduwa5 = as.data.frame(meanValues)
```

## histogram

```{r echo=TRUE}
#ggplot
base1= ggplot(eduwa5,aes(x=meanValues)) 
histogram1= base1 + geom_histogram( binwidth = 2,boundary=0, fill="turquoise3") 
histogram2= histogram1 + stat_bin(binwidth = 2, aes(label=..count..), 
                  geom = "text",boundary = 0,vjust=-0.5) +
                  labs(title="2017 Schools students teacher ratio", 
                       caption = "Source: Seattle Public Schools",
                       x= "Mean Students Teacher ratio",
                       y="Frequency") +
                       xlim(c(10, 24)) +
          scale_x_continuous(breaks = c(10,12,14, 16, 18,20, 22, 24)) +
          geom_vline(aes(xintercept= mean(meanValues), 
                         linetype = "mean"), 
                        colour= 'red') +
          geom_vline(aes(xintercept= median(meanValues), 
                         linetype = "median"), 
                         colour= 'blue') +
          scale_linetype_manual(name = "LEGEND", values = c(2, 2), 
                       guide = guide_legend(override.aes = list(color = c("blue", "red")))) +
          theme(legend.justification =c(lty=0.5,lwd=0.5))
                
histogram2
```
